A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2020-01-31 article-information-2019.tex
131am

Here, the two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data. Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition. $L_1$ regularization is applied to both the lowest projection layer and to the $n_k$ subnetworks to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof. The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size. Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times. For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft} and Appendices  \ref{a_ssec:not} and \ref{a_ssec:xnn}.

@@

PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}. ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}. PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}. For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.

@@

Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$. Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}. Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Sections \ref{sec:m_and_m} and \ref{sec:res}. Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}. Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}. For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.

@@

% XNN image

>>>>

Two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data. Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition. $L_1$ regularization is applied to both the lowest projection layer and to the $n_k$ subnetworks to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof. The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size. Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times. For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft} and Appendices  \ref{a_ssec:not} and \ref{a_ssec:xnn}.

@@

PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j$ \cite{esl}. ICE plots are a newer method that describes the local behavior of $g$ with regard to a single instance of an input feature, $x_j$, and for it's corresponding single instance $\mathbf{x}$. PD and ICE can be overlaid in the same plot to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}. Combining PD and ICE can also be indicative of modeled interactions in $g$ or of known weaknesses in PD estimation (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}) when $\text{PD}(X_j, g)$ and $\text{ICE}(x_j, \mathbf{x}, g)$ diverge. For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.

@@

Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features. Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}. Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Sections \ref{sec:m_and_m} and \ref{sec:res}. Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}. Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}. For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=8cm]{img/xnn_arch.png}
	\caption{A diagram of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}$ predictions. Figure adapted from \citet{wf_xnn}.}
	\label{fig:xnn_arch}
\end{figure}
